Research Notes

### 2018-08-02 writeups/notes.tex
writeup typo fix, shorter timestepper for all

Consider the linearized system of equations (to examine the excited perturbation
in the linear legime)
\begin{subequations}\label{se:vol_drive}
    \begin{align}
        \vec{\nabla} \cdot \vec{u} &= 0,\\
        \pd{\vec{u}_1}{t} + \frac{\vec{\nabla}P}{\rho_0}
            + \frac{\rho_1 \vec{g}}{\rho_0}
            &= 0,\\
        \pd{\rho_1}{t} - \frac{u_{1z} \rho_0}{H}
            &= F(x, z, t),\\
        F(x, z, t) &= \Re Fe^{-\frac{(z - z_0)^2}{2\sigma^2}}
            e^{i\p*{k_xx - \omega t}}.
    \end{align}
\end{subequations}
For the linearized system, we can immediately substitute $\pd{}{t} \to -i\omega,
\pd{}{x} \to ik_x$, giving
\begin{align*}
    \pd{u_{1z}}{z} + ik_xu_x &= 0,\\
    -i\omega u_x + \frac{ik_xP}{\rho_0} &= 0,\\
    -i\omega u_{1z} + \frac{1}{\rho_0}\pd{P}{z} + \frac{\rho_1 g}{\rho_0} &= 0,\\
    -i\omega \rho_1 - \frac{u_{1z}\rho_0}{H} &= F.
\end{align*}
This can be recast solely in terms of $u_{1z}$ as
\begin{align*}
    -i\omega u_{1z} + \frac{\omega}{k_x}\p*{-\frac{1}{ik_x}\ptd{u_{1z}}{z}
        + \frac{1}{ik_xH}\pd{u_{1z}}{z}} - \frac{u_{1z}g}{Hi\omega}
        - \frac{Fg}{i\omega \rho_0} = 0,\\
    \omega^2\p*{
        k_x^2u_{1z} + \frac{1}{H}\pd{u_{1z}}{z} - \ptd{u_{1z}}{z}
    } - u_{1z}N^2k_x^2 - \frac{Fgk_x^2}{\rho_0} = 0.
\end{align*}
Note that if $\pd{}{z} \to 1/2H, F = 0$ then this simplifies to the expected
dispersion relation $\omega^2 = \frac{N^2k_x^2}{k_x^2 + k_z^2 +
\frac{1}{4H^2}}$.

If we approximate $F(x, z, t) \approx F_0\delta\p*{z - z_0} e^{i(k_xx - \omega t)}$
for the time being, then clearly the only term that can be equally singluar is
the $\ptd{}{z}$ term, leaving us with
\begin{equation}
    -\omega^2 \ptd{u_{1z}}{z} - \frac{Fgk_x^2}{\rho_0} = 0.
\end{equation}
Since $u_{1z}$ both above and below $z_0$ satisfy the homogeneous PDE, which has
plane wave solutions with an $e^{z/2H}$ envelope where $k_zH \gg 1$, we can
integrate about a small domain around $z_0$ and obtain
\begin{equation}
    \at{\pd{u_{1z}}{z}}_{z > z_0} - \at{\pd{u_{1z}}{z}}_{z < z_0}
        = -F_0\frac{gk_x^2}{\rho_0\omega^2}e^{i(k_xx - \omega t)}.
\end{equation}
It should then be apparent that the $u_{1z}$ solution is
\begin{equation}
    u_{1z} = \frac{F_0gk_x^2}{\rho_0(z = z_0)\omega^2}\frac{1}{2ik_z} \times
    \begin{cases}
        -e^{(z - z_0)/2H}e^{i(k_xx + k_z(z - z_0) - \omega t)} & z > z_0\\
        e^{(z - z_0)/2H}e^{i(k_xx + k_z(z - z_0) - \omega t)} & z < z_0\\
    \end{cases}.
\end{equation}

>>>>

Consider the linearized system of equations (to examine the excited perturbation
in the linear legime)
\begin{subequations}\label{se:vol_drive}
    \begin{align}
        \vec{\nabla} \cdot \vec{u} &= 0,\\
        \pd{\vec{u}_1}{t} + \frac{\vec{\nabla}P}{\rho_0}
            + \frac{\rho_1 \vec{g}}{\rho_0}
            &= 0,\\
        \pd{\rho_1}{t} - \frac{u_{1z} \rho_0}{H}
            &= F(x, z, t),\\
        F(x, z, t) &= \Re Fe^{-\frac{(z - z_0)^2}{2\sigma^2}}
            e^{i\p*{k_xx - \omega t}}.
    \end{align}
\end{subequations}
For the linearized system, we can immediately substitute $\pd{}{t} \to -i\omega,
\pd{}{x} \to ik_x$, giving
\begin{align*}
    \pd{u_{1z}}{z} + ik_xu_x &= 0,\\
    -i\omega u_x + \frac{ik_xP}{\rho_0} &= 0,\\
    -i\omega u_{1z} + \frac{1}{\rho_0}\pd{P}{z} + \frac{\rho_1 g}{\rho_0} &= 0,\\
    -i\omega \rho_1 - \frac{u_{1z}\rho_0}{H} &= F.
\end{align*}
This can be recast solely in terms of $u_{1z}$ as
\begin{align*}
    -i\omega u_{1z} + \frac{\omega}{k_x}\p*{-\frac{1}{ik_x}\ptd{u_{1z}}{z}
        + \frac{1}{ik_xH}\pd{u_{1z}}{z}} - \frac{u_{1z}g}{Hi\omega}
        - \frac{Fg}{i\omega \rho_0} = 0,\\
    \omega^2\p*{
        k_x^2u_{1z} + \frac{1}{H}\pd{u_{1z}}{z} - \ptd{u_{1z}}{z}
    } - u_{1z}N^2k_x^2 - \frac{Fgk_x^2}{\rho_0} = 0.
\end{align*}
Note that if $\pd{}{z} \to 1/2H + ik_z, F = 0$ then this simplifies to the expected
dispersion relation $\omega^2 = \frac{N^2k_x^2}{k_x^2 + k_z^2 +
\frac{1}{4H^2}}$.

If we approximate $F(x, z, t) \approx F_0\delta\p*{z - z_0} e^{i(k_xx - \omega t)}$
for the time being, then clearly the only term that can be equally singular is
the $\ptd{}{z}$ term, leaving us with
\begin{equation}
    -\omega^2 \ptd{u_{1z}}{z} - \frac{Fgk_x^2}{\rho_0} = 0.
\end{equation}
Since $u_{1z}$ both above and below $z_0$ satisfy the homogeneous PDE, which has
plane wave solutions with an $e^{z/2H}$ envelope where $k_zH \gg 1$, we can
integrate about a small domain around $z_0$ and obtain
\begin{equation}
    \at{\pd{u_{1z}}{z}}_{z > z_0} - \at{\pd{u_{1z}}{z}}_{z < z_0}
        = -F_0\frac{gk_x^2}{\rho_0\omega^2}e^{i(k_xx - \omega t)}.
\end{equation}
It should then be apparent that the $u_{1z}$ solution is
\begin{equation}
    u_{1z} = \frac{F_0gk_x^2}{\rho_0(z = z_0)\omega^2}\frac{1}{2ik_z} \times
    \begin{cases}
        -e^{(z - z_0)/2H}e^{i(k_xx + k_z(z - z_0) - \omega t)} & z > z_0\\
        e^{(z - z_0)/2H}e^{i(k_xx + k_z(z - z_0) - \omega t)} & z < z_0\\
    \end{cases}.
\end{equation}
